Locally-Oriented Programming:\\ A Simple Programming Model for Stencil-Based Computations on Multi-Level Distributed Memory Architectures

### 2015-02-08 papers/LOPe/transformations.tex
A few minor changes before finishing up.

This section describes how LOPe extensions to Fortran have been implemented.  It also describes how
source-to-source transformations have been implemented via rewrite rules and strategies in
StrategoXT.

\subsection{Source-to-Source Transformation Infrastructure}

1. Parsing to produce an AST nodes representated as ATerms.
2. Transformations on the AST nodes represented as rewrite rules and strategies in Stratego/XT
3. Pretty printing of transformed nodes using the pretty printing facilites of BoxLib.

@@

For this work we have developed rewrite strategies and rules in Stratego/XT to rewrite Fortran AST
nodes to C AST nodes (extended with necessary OpenCL keywords).  The C AST ATerms have mostly a
one-to-one correspondence with Fortran terms.  A \texttt{CONCURRENT} procedure is transformed to a
OpenCL kernel (with the \texttt{\_\_kernel} attribute), Fortran formal parameters are transformed
to C parameters (with a direct mapping of types), and local Fortran variable declarations are rewritten
as C declarations.  Similarly, Fortran executable statements are rewritten as C executable AST nodes.
Mostly the only complication is mapping the LOPe local array index view to the global C index space.
This translation is facilitated by the Fortran symbol table that stores array shape and halo
size information.

>>>>

This section briefly describes how LOPe extensions to Fortran have been implemented as
source-to-source transformations via rewrite rules and strategies in StrategoXT.

Transformations are accomplished in three phases:
1. Parsing to produce an AST nodes representated as ATerms;
2. Transformations on the AST nodes represented as rewrite rules and strategies in Stratego/XT; and
3. Pretty printing of transformed nodes using the pretty printing facilites of BoxLib.

@@

For this work we have developed rewrite strategies and rules in Stratego/XT to rewrite Fortran AST
nodes to C AST nodes (extended with necessary OpenCL keywords).  The C AST ATerms have mostly a
one-to-one correspondence with Fortran terms.  A \texttt{CONCURRENT} procedure is transformed to an
OpenCL kernel, Fortran formal parameters are transformed to C parameters (with a direct mapping of
types), and local Fortran variable declarations are rewritten as C declarations.  Similarly,
Fortran executable statements are rewritten as C executable AST nodes.  Mostly the only
complication is mapping the LOPe local array index view to the global C index space.  This
translation is facilitated by a Fortran symbol table that stores array shape and halo size
information.
